Molecular dynamics simulation of synchronization in driven particles

### 2020-10-22 manuscript/manuscript_ajp_2020.tex
updated section labels, single particle section started, added q1D channel

 We create several model light fields,
 creating a landscape of potential minima and maxima
 that modify the local force on a particle as a function of position
 $\vec{F}_{l,i} = \vec{F}_{l}(\vec{r}_i).$
 The landscape potential $V(\vec{r}) $ are static,
 generating a conservative force $\vec{F}_l= -\nabla V_l(\vec{r})$
 with fixed minima and maxima
 that are periodic or quasi-periodic,
 as described in Sec.~\ref{sec:results}.
 One example of a periodic light field is
 that used in Sec.~\ref{sec:kink}
 where 
 \begin{equation}
   V_l = V_0 \cos{(N_p \pi y / L)}
 \end{equation}
 where $N_p$ are the number of troughs in the landscape,
 and $V_0$ is an adjustable parameter
 to set the depth of the troughs.

 We model particle interactions with
 the Yukawa potential $\vec{F}_{ij} = -\nabla V_{ij}(r_{ij})$,
 \begin{equation}
   V_{ij}(r_{ij}) = \frac{E_0}{r_{ij}} e^{-\kappa r_{ij}},
 \end{equation}
 where $r_{ij} = |\vec{r}_i - \vec{r}_j|$.
 This is a
 screened Coulomb potential
 $E_0=2$ scales strength of repulsion
 where $E_0 = k q_1 q_2$  [CHECK scaling/units].
 $\kappa = 1/R_0$ is the screening parameter,
that screening accounts for the lengthscale at
which many particle
interactions and local environment
that reduces the interaction range of individual particles.
We fix $R_0$ to be unity in our simulation units.

We model the 
particle dynamics 
with an overdamped
equation of motion
integrated with the Verlet method.
Here 
the suspending fluid is highly viscous
and exerts a damping force on the particles equivalent
to a linear drag force.
%Overdamped dynamics is a common assumption in models of colloidal particles.
%This is an assumption that 
The damping comes from a viscous fluid model
providing a nonconservative force,
modeled as a linear friction $F_{drag} = -\eta \vec{v}_i$
sufficient so that
the acceleration of the particle is zero.

A single particle has the equation of motion 
\begin{equation}
\eta \vec{v}_i = \vec{F}_{l,i} + \sum_{i \neq j}^{N} \vec{F}_{ij} + \vec{F}_{D}(t).
\end{equation}
where $\eta = 1$.
The equation of motion provides a direct calculation of the velocity
of an individual particle from its location in in space.
Since $\vec{a}$ is zero,
the Verlet method simplifies to 
the Euler method,
which is used to calculate the position. %Most undergraduate students learn the Euler method in Computational Physics and Differential Equations, but don't encounter the Verlet method.
\begin{equation}
\vec{r}_i(t+\Delta t) = \vec{v}_i(t) \Delta t + \vec{r}_i(t).
\end{equation}

@@

We used our model to generate a system, Fig. 1a), where the particle hops between troughs in the energy landscape. In the attached movie, Figure1.mp4, we apply a constant $F_{AC}$ with frequency $f$ then slowly increase $F_{DC}$  to achieve a variety of modes. A mode is a periodic pattern of hops with a constant average particle velocity, $\bar{v}_{y}$ over a range of driving forces $F_{DC}$. In the velocity-force plot in Fig. 1b) $\bar{v}_{y}$ is increasing non-uniformly, with steps, indicating mode locking. A related experiment \cite{juniper2015} revealed the microscopic dynamics of mode locking by driving colloids across a periodic landscape generated by optical tweezers.

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[scale=.25]{single}
\caption{\textbf{(a)} The particle is driven with a constant $F_{AC}$ and $f$ through a periodic potential landscape where blue are minima and red are maxima in the potential. \textbf{(b)} An average particle velocity in the y-direction $\bar{v}_{y}$ as a function of the driving force $F_{DC}$. In the animation the magenta dot represents the $\bar{v}_{y}$ at which the particle in Fig. 1a) is moving.}
\end{figure}
\end{center}

>>>>

 We create several model light fields,
 creating a landscape of potential minima and maxima
 that modify the local force on a particle as a function of position
 $\vec{F}_{l,i} = \vec{F}_{l}(\vec{r}_i).$
 The landscape potential $V(\vec{r}) $ are static,
 generating a conservative force $\vec{F}_l= -\nabla V_l(\vec{r})$
 with fixed minima and maxima
 that are periodic or quasi-periodic,
 as described in Sec.~\ref{sec:results}.
 In every simulation, we confine
 the particles to a quasi-one-dimensional geometry
 using 
 \begin{equation}
   V_{q1D} = V_0 \cos{(\pi x / L)}
 \end{equation}.

 One example of a periodic light field is
 %that used in Sec.~\ref{sec:one}
 where 
 \begin{equation}
   V_l = V_0 \cos{(N_p \pi y / L)}
 \end{equation}
 where $N_p$ are the number of troughs in the landscape,
 and $V_0$ is an adjustable parameter
 to set the depth of the troughs.

 We model particle interactions with
 the Yukawa potential $\vec{F}_{ij} = -\nabla V_{ij}(r_{ij})$,
 \begin{equation}
   V_{ij}(r_{ij}) = \frac{E_0}{r_{ij}} e^{-\kappa r_{ij}},
 \end{equation}
 where $r_{ij} = |\vec{r}_i - \vec{r}_j|$.
 This is a
 screened Coulomb potential
 $E_0=2$ scales strength of repulsion
 where $E_0 = k q_1 q_2$  [CHECK scaling/units].
 $\kappa = 1/R_0$ is the screening parameter 
that accounts for the lengthscale at
which many particle
interactions and local environment
that reduces the interaction range of individual particles.
We fix $R_0$ to be unity in our simulation units.

We model the 
particle dynamics 
with an overdamped
equation of motion
integrated with the Verlet method.
Here 
the suspending fluid is highly viscous
and exerts a damping force on the particles equivalent
to a linear drag force.
%Overdamped dynamics is a common assumption in models of colloidal particles.
%This is an assumption that 
The damping comes from a viscous fluid model
providing a nonconservative force,
modeled as a linear friction $F_{drag} = -\eta \vec{v}_i$
sufficient so that
the acceleration of the particle is zero.
[This model is encountered in Ch. 2 of Taylor's Classical Mechanics....]

A single particle has the equation of motion 
\begin{equation}
\eta \vec{v}_i = \vec{F}_{l,i} + \sum_{i \neq j}^{N} \vec{F}_{ij} + \vec{F}_{D}(t).
\end{equation}
where $\eta = 1$.
The equation of motion provides a direct calculation of the velocity
of an individual particle from its location.
Since we model the acceleration $\vec{a}$ as zero,
the Verlet method simplifies to 
the Euler method,
which is used to calculate the position. %Most undergraduate students learn the Euler method in Computational Physics and Differential Equations, but don't encounter the Verlet method.
\begin{equation}
\vec{r}_i(t+\Delta t) = \vec{v}_i(t) \Delta t + \vec{r}_i(t).
\end{equation}

@@

We place a single particle in a periodic light field is
\begin{equation}
  V_l = V_{0x} \cos{(\pi x/L)} + V_{0y} \cos{(N_p \pi y / L)} 
\end{equation}
where $N_p=20$ are the number of troughs in the landscape,
and $V_{0x}=2$ and $V_{0x}=2$.

We used our model to generate a system, Fig.~\ref{fig:1}(a), where the particle hops between troughs in the energy landscape. In the attached movie, Figure1.mp4, we apply a constant $F_{AC}$ with frequency $f$ then slowly increase $F_{DC}$  to achieve a variety of modes. A mode is a periodic pattern of hops with a constant average particle velocity, $\bar{v}_{y}$ over a range of driving forces $F_{DC}$. In the velocity-force plot in Fig. 1b) $\bar{v}_{y}$ is increasing non-uniformly, with steps, indicating mode locking. A related experiment \cite{juniper2015} revealed the microscopic dynamics of mode locking by driving colloids across a periodic landscape generated by optical tweezers.

\begin{center}
\begin{figure}[h!]
\centering
\includegraphics[scale=.25]{single}
\caption{\textbf{(a)} The particle is driven with a constant $F_{AC}$ and $f$ through a periodic potential landscape where blue are minima and red are maxima in the potential. \textbf{(b)} An average particle velocity in the y-direction $\bar{v}_{y}$ as a function of the driving force $F_{DC}$. In the animation the magenta dot represents the $\bar{v}_{y}$ at which the particle in Fig. 1a) is moving.}
\label{fig:1}
\end{figure}
\end{center}
